
## Setup


#Model parameter/complexity score tests
```{r}
library(tidyverse)  # read_csv and mutate
library(lme4)       # lmer
library(emmeans)    # emmeans and pairs
library(lmerTest)
library(ordinal)
library(performance)
library(modelsummary)
library(lmerTest)

df_wing <- read_csv('/Users/sm6511/Desktop/Prediction-Accomodation-Exp/data/Study2.0/Cleaned/df_wing_for_R-Study2.csv') |>
  mutate(
    task = relevel(factor(task), ref = "accommodate"),
    wing_relevant = relevel(factor(wing_relevant), ref = "1"),
    wing_high = relevel(factor(wing_high), ref = "1")
  )
```

```{r}
df_wing_irrelevant <- df_wing |>
  filter(wing_relevant == 0) |>
  mutate(
    reported_relevant = if_else(feature_importance != 0, 1, 0),
    task = factor(task),
    wing_high = factor(wing_high)
  )

contrasts(df_wing_irrelevant$task) <- contr.sum(2)
contrasts(df_wing_irrelevant$wing_high) <- contr.sum(2)

model_overfit <- glm(
  reported_relevant ~ task * wing_high,
  data = df_wing_irrelevant,
  family = binomial
)

summary(model_overfit)
```

```{r}
df_params <- read_csv('/Users/sm6511/Desktop/Prediction-Accomodation-Exp/data/Study2.0/Cleaned/df_params_study2_for_r.csv')

df_params <- df_params %>%
  mutate(
    task              = factor(task),
    feature_relevance = factor(intuitive_relevance, levels = c(0, 1),
                                labels = c("irrelevant", "relevant")),
    feature_direction = factor(intuitive_direction, levels = c(0, 1),
                                labels = c("no_wing", "wing")),
    overfit           = factor(overfit, levels = c(0, 1),
                                labels = c("no_overfit", "overfit")),
    participant       = factor(participant)
  )

contrasts(df_params$task)              <- contr.sum(2)
contrasts(df_params$feature_relevance) <- contr.sum(2)
contrasts(df_params$feature_direction) <- contr.sum(2)

model_overfit <- glmer(
  overfit ~ task * feature_relevance * feature_direction +
    (1 | participant),
  data    = df_params,
  family  = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

summary(model_overfit)
```


#Importance score test #1 (including direction)
```{r}
contrasts(df_wing$task) <- contr.sum(2)
contrasts(df_wing$wing_relevant) <- contr.sum(2)
contrasts(df_wing$wing_high) <- contr.sum(2)

model_wing <- lm(
  feature_importance ~ task * wing_relevant * wing_high,
  data = df_wing
)

summary(model_wing)
```

```{r}
emm <- emmeans(
  model_wing,
  ~ task * wing_relevant * wing_high
)

emm
pairs(
  emm,
  by = c("wing_relevant", "wing_high")
)
```

#Importance score test #2 (all features, no direction)

```{r}
library(tidyverse)  # read_csv and mutate
library(lme4)       # lmer
library(emmeans)    # emmeans and pairs
library(lmerTest)
library(ordinal)
library(performance)
library(modelsummary)
library(lmerTest)

df_long4 <- read_csv('/Users/sm6511/Desktop/Prediction-Accomodation-Exp/data/Study2.0/Cleaned/df_long_for_R-Study2.csv') |>
  mutate(
    task = relevel(factor(task), ref = "predict"),
    feature_relevance = relevel(factor(feature_relevance), ref = "irrelevant"),
    feature_dimension = relevel(factor(feature_dimension), ref = "color"),
  )
```

```{r}
#| fig-width: 14
#| fig-height: 10
contrasts(df_long4$task) <- contr.sum(2)
contrasts(df_long4$feature_relevance) <- contr.sum(2)
contrasts(df_long4$feature_dimension) <- contr.sum(3)

model_eff <- lmer(
  feature_importance ~ task * feature_relevance *
    feature_dimension +
    (1|participant),
  data = df_long4
)
summary(model_eff)
```

#Followup tests
```{r}
joint_tests(model_eff)
emm_task_by_dimension <- emmeans(
  model_eff,
  ~ task | feature_dimension
)

pairs(
  emm_task_by_dimension,
  adjust = "holm"
)
```

```{r}
levels(df_long4$task)
levels(df_long4$feature_dimension)
contrasts(df_long4$task)
contrasts(df_long4$feature_dimension)
emm_2way <- emmeans(model_eff, ~ task * feature_dimension)
emm_2way
contrast(emm_2way, interaction = c("pairwise", "pairwise"))
emmip(model_eff, task ~ feature_dimension, CIs = TRUE)
```


```{r}
library(tidyverse)

plot_df <- df_long4 %>%
  group_by(task, feature_relevance, feature_dimension) %>%
  summarise(
    mean = mean(feature_importance, na.rm = TRUE),
    se = sd(feature_importance, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(
  plot_df,
  aes(
    x = feature_dimension,
    y = mean,
    color = task,
    group = task
  )
) +
  geom_point(
    position = position_dodge(width = .3),
    size = 3
  ) +
  geom_line(
    position = position_dodge(width = .3)
  ) +
  geom_errorbar(
    aes(
      ymin = mean - se,
      ymax = mean + se
    ),
    width = .15,
    position = position_dodge(width = .3)
  ) +
  facet_wrap(~ feature_relevance) +
  theme_classic() +
  labs(
    x = "Feature Dimension",
    y = "Mean Feature Importance",
    color = "Task"
  )
```


```{r}
emm <- emmeans(
  model_eff,
  ~ task * feature_relevance * feature_dimension
)

emm
pairs(
  emm,
  by = c("feature_relevance", "feature_dimension")
)
```

#Prereg wing direction correlation analysis
```{r}
df_corr <- read_csv('/Users/sm6511/Desktop/Prediction-Accomodation-Exp/data/Combined/df_corr_for_R_Study2.csv') |>
  mutate(
    task = factor(task),
    feature_relevance = factor(feature_relevance),
    feature_dimension = factor(feature_dimension)
  )

df_wing_corr <- df_corr %>%
  filter(feature_dimension == "wing") %>%
  mutate(
    wing_relevant = factor(irrelevant_dim != "wing"),
    wing_high  = factor(high_value == "T"),
    task          = factor(task)
  ) %>%
  mutate(
    task          = relevel(task, ref = "accommodate"),
    wing_relevant = relevel(wing_relevant, ref = "TRUE"),
    wing_high  = relevel(wing_high, ref = "TRUE")
  )
```

```{r}
contrasts(df_wing_corr$wing_relevant) <- contr.sum(2)
contrasts(df_wing_corr$task) <- contr.sum(2)
contrasts(df_wing_corr$wing_high) <- contr.sum(2)

model_corr_raw_eff <- lm(
  correlation ~ task * wing_relevant * wing_high,
  data = df_wing_corr
)
summary(model_corr_raw_eff)

```


#correlation test 2 (no direction, include all features)
```{r}

# Effect / sum coding
contrasts(df_corr$task) <- contr.sum(2)
contrasts(df_corr$feature_relevance) <- contr.sum(2)
contrasts(df_corr$feature_dimension) <- contr.sum(3)

model_corr_raw_eff <- lmer(
  correlation ~ task * feature_relevance * feature_dimension +
    (1 | participant),
  data = df_corr
)

summary(model_corr_raw_eff)
```

#Followup tests
```{r}
# Estimated marginal means of relevance in each dimension
emm_rel_dim <- emmeans(
  model_corr_raw_eff,
  ~ feature_relevance | feature_dimension
)

# Relevant vs. irrelevant separately for each dimension
pairs(
  emm_rel_dim,
  adjust = "holm"
)
```

```{r}

emm <- emmeans(
  model_corr_raw_eff,
  ~ task * feature_relevance * feature_dimension
)

emm
pairs(
  emm,
  by = c("feature_relevance", "feature_dimension")
)
```


```{r}
library(tidyverse)

plot_df <- df_corr %>%
  group_by(task, feature_relevance, feature_dimension) %>%
  summarise(
    mean = mean(correlation, na.rm = TRUE),
    se = sd(correlation, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(
  plot_df,
  aes(
    x = feature_dimension,
    y = mean,
    color = task,
    group = task
  )
) +
  geom_point(
    position = position_dodge(width = .3),
    size = 3
  ) +
  geom_line(
    position = position_dodge(width = .3)
  ) +
  geom_errorbar(
    aes(
      ymin = mean - se,
      ymax = mean + se
    ),
    width = .15,
    position = position_dodge(width = .3)
  ) +
  facet_wrap(~ feature_relevance) +
  theme_classic() +
  labs(
    x = "Feature Dimension",
    y = "Mean Correlation",
    color = "Task"
  )
```


#Supplement in prereg (look at actual associations)
```{r}

df_irrelevant <- subset(df_corr,
                        feature_relevance == "irrelevant")

model <- lm(
    correlation ~ task * actual_correlation,
    data = df_irrelevant
)
summary(model)
```

```{r}
df_merged <- read_csv('/Users/sm6511/Desktop/Prediction-Accomodation-Exp/data/Combined/df_merged_for_R_Study2.csv') |>
  mutate(
    task = relevel(factor(task), ref = "predict"),

    color_relevance = ifelse(irrelevant_dim == "color", "irrelevant", "relevant"),
    tail_relevance  = ifelse(irrelevant_dim == "tail",  "irrelevant", "relevant"),
    wing_relevance  = ifelse(irrelevant_dim == "wing",  "irrelevant", "relevant"),

    color_match_high = ifelse(color == color_high, 1, 0),
    tail_match_high  = ifelse(tail  == tail_high,  1, 0),
    wing_match_high  = ifelse(wing  == wing_high,  1, 0)
  )
```


###PREREG WING DIRECTIONAL ERROR ANALYSIS

```{r}

df_wing_cue <- df_merged |>
  mutate(
    #First, error for each exemplar
    # Positive = participant predicted more food than observed
    # Negative = participant predicted less food than observed
    error = testing_responses - food_amount,
    abs_error = abs(error),

    # Whether wings are objectively relevant in this condition
    wing_relevance = if_else(
      irrelevant_dim == "wing",
      "irrelevant",
      "relevant"
    ),

    # Intuitive direction:
    # T = wings
    # N = paws/no wings
    intuitive_direction = case_when(
      wing == "T" ~ "wings",
      wing == "N" ~ "legs",
      TRUE ~ NA_character_
    )
  ) |>
  filter(
    !is.na(error),
    !is.na(intuitive_direction),
    !is.na(wing_relevance)
  ) |>
  group_by(
    task,
    conditionOrder,
    relevant_dim_1,
    relevant_dim_2,
    irrelevant_dim,
    wing_relevance,
    intuitive_direction
  ) |>
  summarise(
    signed_error = mean(error, na.rm = TRUE),
    abs_error = mean(abs_error, na.rm = TRUE),
    n_exemplars = sum(!is.na(error)),
    .groups = "drop"
  ) |>
  mutate(
    task = relevel(factor(task), ref = "predict"),

    wing_relevance = factor(
      wing_relevance,
      levels = c("irrelevant", "relevant")
    ),

    intuitive_direction = factor(
      intuitive_direction,
      levels = c("legs", "wings")
    ),

    # A participant identifier unique across conditions/tasks
    participant = interaction(
      task,
      conditionOrder,
      drop = TRUE
    )
  )
```

```{r}
library(lme4)
model_wing_error <- lm(
  abs_error ~
    task *
    wing_relevance *
    intuitive_direction, 
  data = df_wing_cue)

summary(model_wing_error)
```
```{r}
library(tidyverse)

#summary stats
plot_summary <- df_wing_cue |>
  group_by(wing_relevance, intuitive_direction) |>
  summarise(
    mean_signed = mean(signed_error, na.rm = TRUE),
    se_signed = sd(signed_error, na.rm = TRUE) / sqrt(n()),

    mean_abs = mean(abs_error, na.rm = TRUE),
    se_abs = sd(abs_error, na.rm = TRUE) / sqrt(n()),

    .groups = "drop"
  )

# Consistent ordering
plot_summary$wing_relevance <-
  factor(plot_summary$wing_relevance,
         levels = c("irrelevant", "relevant"))

plot_summary$intuitive_direction <-
  factor(plot_summary$intuitive_direction,
         levels = c("legs", "wings"))
ggplot(
  plot_summary,
  aes(
    x = wing_relevance,
    y = mean_signed,
    fill = intuitive_direction
  )
) +
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7
  ) +
  geom_errorbar(
    aes(
      ymin = mean_signed - se_signed,
      ymax = mean_signed + se_signed
    ),
    width = .18,
    position = position_dodge(width = 0.8)
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    x = "Wing relevance",
    y = "Signed error",
    fill = "Intuitive direction"
  ) +
  theme_classic(base_size = 18)
ggplot(
  plot_summary,
  aes(
    x = wing_relevance,
    y = mean_abs,
    fill = intuitive_direction
  )
) +
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7
  ) +
  geom_errorbar(
    aes(
      ymin = mean_abs - se_abs,
      ymax = mean_abs + se_abs
    ),
    width = .18,
    position = position_dodge(width = 0.8)
  ) +
  labs(
    x = "Wing relevance",
    y = "Absolute error",
    fill = "Intuitive direction"
  ) +
  theme_classic(base_size = 18)
```


###REGULAR ERROR ANALYSIS



```{r}
df_cue <- df_merged |>
  pivot_longer(
    cols = c(color_match_high, tail_match_high, wing_match_high),
    names_to = "feature_dimension",
    values_to = "match_high"
  ) |>
  mutate(
    feature_dimension = str_remove(feature_dimension, "_match_high"),
    feature_relevance = case_when(
      feature_dimension == "color" ~ color_relevance,
      feature_dimension == "tail"  ~ tail_relevance,
      feature_dimension == "wing"  ~ wing_relevance
    )
  ) |>
  group_by(
    task, conditionOrder,
    relevant_dim_1, relevant_dim_2, irrelevant_dim,
    feature_dimension, feature_relevance
  ) |>
  summarise(
    pred = mean(testing_responses[match_high == 1], na.rm = TRUE) -
           mean(testing_responses[match_high == 0], na.rm = TRUE),

    true = mean(food_amount[match_high == 1], na.rm = TRUE) -
           mean(food_amount[match_high == 0], na.rm = TRUE),

    .groups = "drop"
  ) |>
  mutate(
    signed_error = pred - true,
    abs_error = abs(pred - true),

    task = relevel(factor(task), ref = "predict"),
    feature_dimension = relevel(factor(feature_dimension), ref = "tail"),
    feature_relevance = relevel(factor(feature_relevance), ref = "irrelevant")
  )
```

```{r}
contrasts(df_cue$feature_dimension)
```
```{r}

contrasts(df_cue$task) <- contr.sum(2)
contrasts(df_cue$feature_relevance) <- contr.sum(2)
contrasts(df_cue$feature_dimension) <- contr.sum(3)

model_error <- lm(
  abs_error ~ task * feature_relevance * feature_dimension,
  data = df_cue
)

summary(model_error)
```

```{r}
joint_tests(model_signed)
```

```{r}
emmeans(model_signed, pairwise ~ feature_relevance | feature_dimension)
```

```{r}
emm <- emmeans(
  model_error,
  ~ task * feature_relevance * feature_dimension
)

emm
pairs(
  emm,
  by = c("feature_relevance", "feature_dimension")
)
```

```{r}
library(ggplot2)

ggplot(
  df_cue,
  aes(
    x = feature_dimension,
    y = abs_error,
    color = task,
    group = task
  )
) +
  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = .2),
    size = 3
  ) +
  stat_summary(
    fun = mean,
    geom = "line",
    position = position_dodge(width = .2)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = .1,
    position = position_dodge(width = .2)
  ) +
  facet_wrap(~feature_relevance) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    x = "Feature",
    y = "Error (Participant effect − True effect)",
    color = "Task"
  ) +
  theme_classic()
```

```{r}
library(ggplot2)

ggplot(
  df_cue,
  aes(
    x = feature_dimension,
    y = abs_error,
    color = feature_relevance,
    group = feature_relevance
  )
) +
  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = .2),
    size = 3
  ) +
  stat_summary(
    fun = mean,
    geom = "line",
    position = position_dodge(width = .2)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = .1,
    position = position_dodge(width = .2)
  ) +
  facet_wrap(~feature_relevance) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    x = "Feature",
    y = "Error (Participant effect − True effect)",
    color = "Relevance"
  ) +
  theme_classic()
```


